---
title: "RepReserach-PA1"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reproducible Research: Peer Assessment 1

library(ggplot2)
library(scales)
library(Hmisc)

## Loading and preprocessing the data
### 1. Load the data (i.e. read.csv())

```{r load data, echo=TRUE}
setwd('C:/Users/Home/Downloads/Coursera-DS/Course5-ReproducibleResearch/Lesson2')

if(!file.exists('activity.csv')){
    unzip('activity.zip')
}
actdata <- read.csv('activity.csv')
```

### 2. Process/transform the data (if necessary) into a format suitable for your analysis

```{r optional transform, echo=TRUE}
#actdata$interval <- strptime(gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", actdata$interval), format='%H:%M')
```

## What is mean total number of steps taken per day?


```{r steps, echo=TRUE}
stepsByDay <- tapply(actdata$steps, actdata$date, sum, na.rm=TRUE)
```

### 1. Make a histogram of the total number of steps taken each day

```{r plotsteps, echo=TRUE}
library(ggplot2)
qplot(stepsByDay, xlab='Total steps per day', ylab='Frequency using binwith 500', binwidth=500)
```

### 2. Calculate and report the mean and median total number of steps taken per day

```{r mean median, echo=TRUE}
stepsByDayMean <- mean(stepsByDay)
stepsByDayMedian <- median(stepsByDay)

#  Mean: r 
stepsByDayMean
#  Median: r 
stepsByDayMedian
```

## What is the average daily activity pattern?

```{r meansteps, echo=TRUE}
averageStepsPerTimeBlock <- aggregate(x=list(meanSteps=actdata$steps), by=list(interval=actdata$interval), FUN=mean, na.rm=TRUE)
```

1. Make a time series plot

```{r meansteps plot, echo=TRUE}
ggplot(data=averageStepsPerTimeBlock, aes(x=interval, y=meanSteps)) +
    geom_line() +
    xlab("5-minute interval") +
    ylab("average number of steps taken") 
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r max steps, echo=TRUE}
mostSteps <- which.max(averageStepsPerTimeBlock$meanSteps)
timeMostSteps <-  gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", averageStepsPerTimeBlock[mostSteps,'interval'])
```
Most Steps at: r 
 ```{r moststeps, echo=TRUE}
timeMostSteps
```

## Imputing missing values
1. Calculate and report the total number of missing values in the dataset

```{r missing values, echo=TRUE}
numMissingValues <- length(which(is.na(actdata$steps)))
```
    Number of missing values: r 

```{r report no. of missing values, echo=TRUE}    
    numMissingValues
```

2. Devise a strategy for filling in all of the missing values in the dataset.

Substitute the missing values with the mean values of steps. This is done using
the impute function in Hmisc library.

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.


```{r impute missing values, echo=TRUE} 
library(Hmisc)
actdataImputed <- actdata
actdataImputed$steps <- impute(actdata$steps, fun=mean)
```

4. Make a histogram of the total number of steps taken each day

```{r histogram for missing values, echo=TRUE} 
stepsByDayImputed <- tapply(actdataImputed$steps, actdataImputed$date, sum)
qplot(stepsByDayImputed, xlab='Total steps per day (Imputed)', ylab='Frequency using binwith 500', binwidth=500)
```

... and Calculate and report the mean and median total number of steps taken per day.

```{r mean median imputed, echo=TRUE} 
stepsByDayMeanImputed <- mean(stepsByDayImputed)
stepsByDayMedianImputed <- median(stepsByDayImputed)
```
    Mean (Imputed): r 
```{r print imputed mean, echo=TRUE} 
    stepsByDayMeanImputed
```

    Median (Imputed): r 
```{r print imputed median, echo=TRUE}     
    stepsByDayMedianImputed
```

    ## Are there differences in activity patterns between weekdays and weekends?
1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r date type, echo=TRUE} 
actdataImputed$dateType <-  ifelse(as.POSIXlt(actdataImputed$date)$wday %in% c(0,6), 'weekend', 'weekday')
```

2. Make a panel plot containing a time series plot

```{r time series panel plot, echo=TRUE} 
averagedactdataImputed <- aggregate(steps ~ interval + dateType, data=actdataImputed, mean)
ggplot(averagedactdataImputed, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(dateType ~ .) +
    xlab("5-minute interval") + 
    ylab("avarage number of steps")


